Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-12 paper/sections/proof-reconstruction.tex
[ paper ] complexity measure added.

\subsection{Rules as Theorems}

Main algorithms in \CPL need the propositions in some of their normal
forms. For instance, from the negative normal form
(henceforth, \abbre{NNF}), we can derive the conjunctive normal form,
and the disjunctive normal form as well.
Therefore, we start giving the definition of \abbre{NNF}, one
function definition, \name{nnf} and a refinement of it.
We will use it later for instance in the description of \canonicalize
rule. The following treatment to this function, \name{nnf},
applies for other conversion functions and theorems defined in
\cite{AgdaProp,AgdaMetis}.
We use the notation convention $:=$ for equivalence between
propositions. The \PROP inductive data type is defined as usual
based on the definition from section \ref{sec:preliminaries} and
implemented in \Agda with \name{PropFormula} in section
\ref{sec:implementation}.

\subsubsection{\CPL Functions}
\hspace*{5cm}\\[3mm]
\emph{NNF.}
A negative normal form for the formula φ is one in which negations
appear only in the literals and the expression is in terms only of
the boolean operators (¬, ∧, ∨).

To transform a formula φ to its equivalent \abbre{NNF} form, we
apply recursively the following equations. The function type for
\name{nnf} is $\PROP \to \PROP$.

\begin{equation}
\label{eq:nnf-definition}
\nnf(φ) =
\begin{cases}
\nnf(φ₁) ∧ \nnf(φ₂)
   & φ := φ₁ ∧ φ₂  \\

\nnf(φ₁) ∨ \nnf(φ₂)
   & φ := φ₁ ∨ φ₂  \\

\nnf(¬ φ₁ ∨ φ₂)
   & φ := φ₁ ⇒ φ₂  \\

\nnf(φ₁ ⇒ φ₂ ∧ φ₂ ⇒ φ₁)
   & φ := φ₁ ⇔ φ₂\\

\nnf(¬ φ₁ ∨ ¬ φ₂)
   & φ := ¬ (φ₁ ∧ φ₂) \\

\nnf(¬ φ₁ ∧ ¬ φ₂)
   & φ := ¬ (φ₁ ∨ φ₂) \\

\nnf(φ₁)
   & φ := ¬ ¬ φ₁     \\

\nnf(¬ (φ₂ ∨ ¬ φ₁))
   & φ := ¬ (φ₁ ⇒ φ₂)  \\

\nnf(¬ (φ₁ ⇒ φ₂ ∧ φ₂ ⇒ φ₁))
   & φ := ¬ (φ₁ ⇔ φ₂) \\

⊥  & φ := ¬ ⊤        \\

⊤  & φ := ¬ ⊥        \\

φ  & \text{otherwise}
\end{cases}
\end{equation}

A recursive function is a well-structured recursion, if the size of
the arguments with which it is defined decrease in size for each
call to the function in the definition. That guarantee termination
of the function, and its translation to constructive Type Theory.

But we need to improve the definition for \name{nnf} in order to have
a version of a well-structured recursion.
We show a strategy to achieve that with a smaller example function,
\unshunt, that we will use forward to formalize the inference
\name{strip}.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt(φ) =
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃)\\

\unshunt(φ₁ ⇒ φ₂) ∧ \unshunt(φ₁ ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃)\\

φ &\text{otherwise}
\end{cases}
\end{equation}

To translate the functions in a way that is well defined,
we used the method proposed by Bove and Capettra in
\cite{Bove2005}, which basically consists of

We take a general recursive function $f: \sigma \to \tau$.
Its formalization in Type Theory will define an inductive
predicate $Ind$, $ P: \sigma \to Ind$.

In our case, the set of natural numbers will serve as our
inductive set $Ind$ and as second argument for new function.
Thus, the function $\unshunt$ will be defined by structural
recursion in his second term.

The new proposed $\unshunt^{*}$ has type,

\[
\unshunt^{*} : PROP \to \mathbb{N} \to PROP.
\]

\begin{equation}
\label{eq:unshunt-structured}
\unshunt^{*}(φ, n) =
\begin{cases}
\unshunt^{*}((φ₁ ∧ φ₂) ⇒ φ₃, n-1)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃)\\

\unshunt^{*}(φ₁ ⇒ φ₂, n-1) ∧ \unshunt^{*}(φ₁ ⇒ φ₃,n-1)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃)\\

φ & n := 0
\end{cases}
\end{equation}

\begin{definition}
Th rank of proposition φ denoted by $rank(\phi)$ is defined by the
following equation.

\begin{equation}
\label{eq:rank-definition}
rank(\phi)=
\begin{cases}
0 &\text{for atomic }\phi\\

\max(rank(\phi_1),rank(\phi_2)) + 1
  & \phi := \phi_1\,\square\, \phi_2 \\
rank(\phi_1) + 1  & \phi := \neg \phi_1
\end{cases}
\end{equation}
\end{definition}

The final review for $\unshunt$ function is the
$\unshunt^{+}$ function in Eq.~\ref{eq:unshunt-fixed}
with $PROP \to PROP$ type.

\begin{equation}
\label{eq:unshunt-fixed}
\unshunt^{+}(\varphi) = \unshunt^{*}(\varphi, rank(\varphi))
\end{equation}

In the following section, our task will consist of searching
appropriate functions like $rank$ to assure the termination for
transformation functions like $nnf$.

@@

\subsection{Emulation \Metis' Inference Rules}

@@

\begin{theorem}[{\tt thm-$\unshunt^{*}$}] $Γ ⊢ \unshunt^{*}(φ, n)$ for $n\in \mathbb{N}$, then $Γ ⊢ φ$.
\end{theorem}
\begin{proof} Use induction on the cases defined by the result of
$\unshunt^{*}$. When $n = 0$, by definition in Eq.
\ref{eq:unshunt-structured} then $Γ ⊢ φ$.
For the first case, we prove $Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)$ from
$Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$.
Because $n > 0$, using induction on the second argument and the
theorem $∧⇒$\texttt{-to-}$⇒⇒$ from \cite{AgdaProp}.

@@

The second case proves $Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)$ from
 $Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ∧ φ₃), n)$ and the
theorem $⇒∧⇒$\texttt{-to-}$⇒∧$ from \cite{AgdaProp}
\begin{equation*}
\scalebox{0.95}{
$D_1$ := \begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n) ∧ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
D_2 :=
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂) ∧ \unshunt^{*}(φ₁ ⇒ φ₃)$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$D_1$}
\AxiomC{$D_2$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
\RightLabel{\tt ⇒∧⇒-to-⇒∧}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
\end{bprooftree}
\end{equation*}
\end{proof}

The function \mysplit has the type
\[ \mysplit : \PROP \to \PROP
\]
\begin{equation}
\label{eq:split-definition}
\mysplit(φ) =
\begin{cases}
\unshunt(\mysplit(φ₁)) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂)
  & φ := φ₁ ∧ φ₂\\

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂))
  & φ := φ₁ ∨ φ₂\\

\unshunt(φ₁ ⇒ \mysplit(φ₂))
  & φ := φ₁ ⇒ φ₂\\

\unshunt(φ₁ ⇒ \mysplit(φ₂)) ∧ \unshunt(φ₂ ⇒ \mysplit(φ₁))
  & φ := φ₁ ⇔ φ₂\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & φ := ¬ (φ₁ ∧ φ₂)\\

\unshunt(\mysplit(¬ φ₁)) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂))
  & φ := ¬ (φ₁ ∨ φ₂)\\

\unshunt(\mysplit(φ₁)) ∧ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
  & φ := ¬ (φ₁ ⇒ φ₂)\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))
  & φ := ¬ (φ₁ ⇔ φ₂)\\

\unshunt(\mysplit(φ₁)
  & φ := ¬ ¬ φ₁\\

⊤ & φ := ¬ ⊥\\

⊥ & φ := ¬ ⊤\\

φ &\text{otherwise}
\end{cases}
\end{equation}

\begin{theorem}[Splitting a Goal] $Γ ⊢ \mysplit(φ) ⇒ φ$
\end{theorem}

\begin{proof}
ss
\end{proof}

The \mysplit function returns a conjunction of the subgoal. From each
\TSTP derivation, we have information of the subgoals but no how they strictly appear in that conjunction. For that reason, with \Athena, we infer the exact order in the conjunction that coincides with the output by \mysplit.

>>>>

\subsection{Well-structured Recursion}
\label{ssec:well-structured recursion}

To emulate the \Metis' Inference Rules in type theory.
We encounter many functions involved with termination problems, that
means, it is not possible to determine effectively if the function
terminates. That makes impossible to translate in type theory
directly. For that reason, we show a technique that will help us to fix such a issue and guide us how to proceed later.

A recursive function is a \emph{well-structured recursion}, if the
size of the arguments with which it is defined decrease in
size for each call to the function in the definition.
A well-structured function terminates and it can be translated
directly to type theory.

Consider the following example from the \strip inference.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt(φ) =
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\

\unshunt(φ₁ ⇒ φ₂) ∧ \unshunt(φ₁ ⇒ φ₃)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃),\\

φ &\text{otherwise}.
\end{cases}
\end{equation}

From the above definition, in the first case of the \unshunt
function, the size of the argument doesn't really decrease.
Actually, it maintains with the same size ((φ₁ ∧ φ₂) ⇒ φ₃ and  (φ₁ ⇒
(φ₂ ⇒ φ₃))). Therefore, a termination checking will fail with \unshunt
function.

To improve it, we will use the technique proposed by Bove and Capettra in
\cite{Bove2005}, which in our case basically consists of
take a general recursive function $f: \sigma \to \tau$.
and an predicate $P: \sigma \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structure inductive definition.
Then, we redefine the function as
$f: \sigma \to \abbre{Ind}\ \to \tau$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

In our case, $\abbre{Ind}$ will be the \abbre{Nat} data type for
natural numbers with constructors \abbre{zero} and \abbre{suc} defined as usual.

Thus, we have function $\unshunt : \PROP \to \Nat \to \PROP$ defined by structural recursion in his second term with type as follows,

\begin{align*}
\label{eq:unshunt-structured}
\unshunt(φ, \zero) &= φ\\
\unshunt(φ, \suc(n)) &=
\begin{cases}
\unshunt((φ₁ ∧ φ₂) ⇒ φ₃, n)
  & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\unshunt(φ₁ ⇒ φ₂, n) ∧ \unshunt(φ₁ ⇒ φ₃,n)
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃).
\end{cases}
\end{align*}

Finally, we define the complexity measure to handle the second argument for the \unshunt function.

\newcommand{\rank}{\name{rank}}

A \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017} or
a rank function presented in \cite{VanDalen1994} is a function
$\PROP \to \Nat$ that indents to associate a natural number for a
proposition φ. The following function is a complexity measure.

\begin{equation*}
\label{eq:rank-definition}
\rank(φ)=
\begin{cases}
0 &\text{for atomic }φ, \\
\rank(φ₁) + c_{¬}  & φ := \neg φ₁, \\
\max(\rank(φ₁),\rank(φ₂)) + c_{\square}
  & φ := φ₁\,\square\, φ₂.
\end{cases}
\end{equation*}
where $c_{¬}$ or $c_{\square}$ is a non-negative integer called the
complexity of the connective.

Nevertheless, we need to define an appropriate complexity measure for the second argument in the $\unshunt$ function, in order to
expect the same behavior of the original function as it was defined at first in Eq.~\ref{eq:unshunt-definition}.

\begin{equation}
\label{eq:rank-definition}
\unshunt_{cm}(φ)=
\begin{cases}
\unshunt_{cm}(φ₃) + 2 & φ := φ₁ ⇒ (φ₂ ⇒ φ₃),\\
\max(\unshunt_{cm}(φ₂),\unshunt_{cm}(φ₃)) + 1
  & φ := φ₁ ⇒ (φ₂ ∧ φ₃),\\
0 &\text{otherwise.}
\end{cases}
\end{equation}

Then, a curated version of \unshunt is presented in Eq.
\ref{eq:unshunt-fixed}.

\begin{equation}
\label{eq:unshunt-fixed}
\unshunt(\varphi) := \unshunt(\varphi, \unshunt_{cm}(\varphi))
\end{equation}

In the following section, we may need define such complexity measure
functions for termination-checking.

@@

\subsection{Emulating \Metis' Inference Rules}
\label{ssec:emulating-inferences}

@@

\begin{theorem}[{\tt thm-$\unshunt^{*}$}] $Γ ⊢ \unshunt^{*}(φ, n)$
for $n\in \Nat$, then $Γ ⊢ φ$.
\label{thm:unshunt}
\end{theorem}

\begin{proof} Use induction on the cases defined by the result of
$\unshunt^{*}$. When $n = 0$, by definition in Eq.
\ref{eq:unshunt-structured} then $Γ ⊢ φ$.
For the first case, we prove $Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)$ from
$Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$.
Because $n > 0$, using induction on the second argument and the
theorem $∧⇒$\texttt{-to-}$⇒⇒$ from \cite{AgdaProp}.

@@

The second case proves $Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)$ from
 $Γ ⊢ \unshunt^{*}(φ₁ ⇒ (φ₂ ∧ φ₃), n)$ and the
theorem $⇒∧⇒$\texttt{-to-}$⇒∧$ from \cite{AgdaProp}
\begin{equation*}
\scalebox{0.95}{
$\mathcal{D}_1$ := \begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n) ∧ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
\end{bprooftree}
}
\end{equation*}

\begin{equation*}
\mathcal{D}_2 :=
\begin{bprooftree}
\AxiomC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₂) ∧ \unshunt^{*}(φ₁ ⇒ φ₃)$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n)$}
\RightLabel{By definition}
\UnaryInfC{$Γ ⊢ \unshunt^{*}(φ₁ ⇒ φ₃, n-1)$}
\RightLabel{\tt thm-$\unshunt^{*}$}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\AxiomC{$\mathcal{D}_2$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
\RightLabel{\tt ⇒∧⇒-to-⇒∧}
\UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
\end{bprooftree}
\end{equation*}
\end{proof}

\vskip 3mm

From now on, we redefine the function $\unshunt$ with the
definition of $\unshunt^{+}$ for no overloading the notation forward.

\begin{theorem}[{\tt thm-\unshunt}] $Γ ⊢ \unshunt(φ)$ then $Γ ⊢ φ$.
\end{theorem}
\begin{proof} It follows from Thm.~\ref{thm:unshunt} using the
\rank function as the second argument.
\end{proof}

The \mysplit function yields a conjunction of subformulas that
implies the proposition φ, the argument of the function. This
function has type $\PROP \to \NN \to PROP$, its definition is
 $\mysplit(φ,0) = φ$, and
for all$n\in \NN$, $n > 0$,

\begin{equation}
\label{eq:split-definition}
\mysplit(φ,n+1) =
\begin{cases}
\unshunt(\mysplit(φ₁,n)) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂,n)
  & φ := φ₁ ∧ φ₂,\\

\unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n))
  & φ := φ₁ ∨ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂,n))
  & φ := φ₁ ⇒ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(φ₂,n)) ∧ \unshunt(φ₂ ⇒ \mysplit(φ₁,n))
  & φ := φ₁ ⇔ φ₂,\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂, n))
  & φ := ¬ (φ₁ ∧ φ₂),\\

\unshunt(\mysplit(¬ φ₁,n)) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂,n))
  & φ := ¬ (φ₁ ∨ φ₂),\\

\unshunt(\mysplit(φ₁,n)) ∧ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂,n))
  & φ := ¬ (φ₁ ⇒ φ₂),\\

\unshunt(φ₁ ⇒ \mysplit(¬ φ₂,n)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁,n))
  & φ := ¬ (φ₁ ⇔ φ₂),\\

\unshunt(\mysplit(φ₁, n))
  & φ := ¬ ¬ φ₁,\\

⊤ & φ := ¬ ⊥,\\

⊥ & φ := ¬ ⊤.
\end{cases}
\end{equation}

\newcommand{\thmsplit}{\name{thm-split}}
\begin{theorem}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\end{theorem}

\begin{proof} We use induction on the structure of the formula
following the cases in the definition of \mysplit function.
We proceed to show some cases since the others are proved in a
similar way.\\[3mm]
(φ := φ₁ ∧ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

\AxiomC{$\mathcal{D}_1$}
\UnaryInfC{Γ ⊢ φ₁}

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By Def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt $∧$-proj$₂$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt $⇒$-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ φ₂$}

\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ φ₁ ∧ φ₂$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D}_1 =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ)$}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁) ∧ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁$}
\end{bprooftree}
\end{equation*}

(φ := φ₁ ⇒ φ₂)
\begin{equation*}
\begin{bprooftree}
\AxiomC{}
\RightLabel{\tt assume φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{weaken φ₁}
\UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
\RightLabel{\tt ⇒-intro}
\UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
\end{bprooftree}
\end{equation*}

(φ := ¬ (φ₁ ∨ φ₂))
\begin{equation*}
\begin{bprooftree}
\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

\AxiomC{$\mathcal{D}$}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}

\AxiomC{$Γ ⊢ \mysplit(φ)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂))$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂)$}
\RightLabel{\tt ⇒-elim}
\BinaryInfC{$Γ ⊢ \mysplit(φ₂)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ ⊢ ¬ φ₂$}
\RightLabel{\tt ∧-intro}
\BinaryInfC{$Γ ⊢ ¬ φ₁ ∧ ¬ φ₂$}
\RightLabel{\tt ¬∧¬-to-¬∨}
\UnaryInfC{$Γ ⊢ ¬ (φ₁ ∨ φ₂)$}
\end{bprooftree}
\end{equation*}
where
\begin{equation*}
\mathcal{D} =
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁) ∧ \unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{\tt ∧-proj₁}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₁)$}
\RightLabel{\thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}

(φ := ¬ (φ₁ ⇔ φ₂))
\begin{equation*}
\begin{bprooftree}

% -- right

\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{By def.}
\UnaryInfC{$Γ ⊢ \unshunt(φ₁ ⇒ \mysplit(¬ φ₂)) ∧ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁)) $}
\RightLabel{\tt ∧-proj₂}
\UnaryInfC{$Γ ⊢ \unshunt(¬ φ₂ ⇒ \mysplit(φ₁))$}
\RightLabel{\tt \thmunshunt}
\UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}
\RightLabel{\tt weaken ¬ φ₂}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂ ⇒ \mysplit(φ₁)$}

% -- right

% -- left
\AxiomC{}
\RightLabel{\tt assume (¬ φ₂)}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ ¬ φ₂$}
\RightLabel{\tt ⇒-elim}

\BinaryInfC{$Γ , ¬ φ₂ ⊢ \mysplit(φ₁)$}
\RightLabel{\thmsplit}
\UnaryInfC{$Γ , ¬ φ₂ ⊢ φ₁ $}
\RightLabel{\tt ⇒-intro}
\UnaryInfC{$Γ ⊢ ¬ φ₂ ⇒ φ₁$}
% -- left

% \RightLabel{\tt ∧-intro}
% \BinaryInfC{$Γ ⊢ (φ₁ ⇒ ¬ φ₂) ∧ (¬ φ₂ ⇒ φ₁)$}
% \RightLabel{\tt ⇒¬∧¬⇒-to-¬⇔}
% \UnaryInfC{$Γ ⊢ ¬ (φ₁ ⇔ φ₂)$}
\end{bprooftree}
\end{equation*}

\end{proof}

The \mysplit function returns a conjunction of the subgoals.
From each \TSTP derivation, we actually have information of the
subgoals but it is not explicit how they appear in the conjunction.
For that reason, with \Athena, we infer the exact order in the
conjunction that coincides with the output by \mysplit.
